---
title: "estimating rates of change
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####preamble
Here we are going to look at the evolution of the topics and pick out those that evolve fast and slow

#### libraries
````{r}
library(scales)
library(mgcv)
library(ggplot2)
library(data.table)
library(tidyr)
library(tidyquant)
library(lubridate)
library(colortools)
library(dplyr)
library(broom)
```````
data
````{r}
a<-fread("EEpaperslong.csv", header=TRUE)
a<-as.data.frame(a)
````

get evolution papers
``````{r}
a1<-a%>%
  filter(evolution_paper==1)%>%
  group_by(year, topic_label)%>%
  summarize(N_papers=sum(present05))%>%
  filter(year>=1900 & year<=2010)

papers<-a%>%
  filter(evolution_paper==1)%>%
  group_by(year)%>%
  summarize(N_total_papers=length(paper_id))%>%
  filter(year>=1900 & year<=2010)

a2<-merge(a1, papers, by="year")
a2$prob<-a2$N_papers/a2$N_total_papers
````````
suppose our revolution runs from 1920-1930.  Let's look at the rate of change for each topic from 10 years on either side, ie, 1910-1940.
````{r}
a2$year<-as.numeric(as.character(a2$year))
unique(a2$year)
a3<-a2%>%
filter(year>=1951)%>%
filter(year <1979)
``````
now do a linear regression on the probabilities of each topic

`````{r}
a4<-a3%>%
group_by(topic_label)%>%
do(fitm=lm(prob~year, data=.))
coefs<-tidy(a4, fitm)%>%
filter(term=="year")
coefs$absestimate<-abs(coefs$estimate)
coefs<-coefs%>%arrange(desc(coefs$absestimate))
`````
let's pull out the 5% of topics that are changing fastest. These are most likely to contribute to the revolution
`````{r}
coef1<-coefs%>%
filter(absestimate>=quantile(coefs$absestimate, 0.90))
``````
let's get a list of them
````{r}
revtopics<-coef1$topic_label
``````
let's subset our original data to see how they change
``````{r}
a4<-a3%>%
filter(topic_label%in%revtopics)
``````
plot
`````{r}
coef2<-coef1%>%
  select(topic_label, absestimate)
a4<-merge(a4, coef2, by="topic_label")
a4$topic_label<-reorder(a4$topic_label,desc(a4$absestimate))
levels(a4$topic_label)
ggplot(a4, aes(y=prob, x=year, colour=as.factor(topic_label)))+
geom_point()+
geom_smooth(method="lm", se=FALSE)+
guides(colour=FALSE)+
facet_wrap(~topic_label, ncol=6, scales="free")+
geom_vline(xintercept = 1956, col="red", size=1, linetype="dotted", alpha=0.5)+
geom_vline(xintercept = 1961, col="red", size=1, linetype="dotted", alpha=0.5)+
theme_classic()+
theme(aspect.ratio = 1)+
  ggsave(filename="rapidly evolving topics.pdf", plot = last_plot(),
device = "pdf",
path = "/Users/Armand/Dropbox/MS\ NatEcolEvol", scale = 1,width = 40,height = 20,units = c("cm"))
``````
This plot, then, orders the most rapidly changing variables in the year 1920-1940, in decreasing order of importance. It also shows when the revolution happened. 

So, if there were a revolution in these years, I would say that its due to an increase in seasonal variation, plant community ecology, phenology --- classic communtiy ecology stuff --- at the expense of descriptive botany. 

Of coruse, had I subsetted to just evolution or ecology papers and topics this would be clearer. 