---
title: "fitting gams EE data"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
####preamble
Here we are going to fit, for each topic, a spline to the absolute haldane number as a function of the interval. We are then going to divide it by the interval in order to get the haldane rate. 

#### libraries
````{r}
library(scales)
library(plyr)
library(reshape2)
library(mgcv)
library(ggplot2)
library(data.table)
```````
####data
`````{r}
d<-read.csv("EEpapers_05_res.csv", header=TRUE)
`````
subset to just those midyears which are later than the startyears for each topi
````{r}
z<-read.csv("Topic_starts.csv", header=TRUE)
colnames(z)[3]<-"startyear"
d<-merge(d, z, by.x="topic")
d<-subset(d, midyear>startyear)
````

```{r}
d.2 <- ddply(d, .(topic), summarise, N = length(abs.hald.num))
min(d.2$N)
```
specify the data for the splines.  The x is interval, y is the abs.hald.number, grouping variable is topic
````{r}
y<-d$abs.hald.num
dat<-as.data.frame(y)
dat$x<-d$interval
dat$topic<-d$topic
``````
find out how many samples we now have per topic

````{r}
checkall<- ddply(dat, .(topic), summarise,
length.haldests =length(y),
length.intervals =length(unique(x)))
`````
Remove all topics with few data points.  There is only one, topic 94 which has 3 intervals. Many of the topics which we are interested in include molecular topics which have very few intervals. Therefore the minimum is set such that they are still present.
````{r}
dat<-subset(dat, !topic=="topic94")
``````
Since some of our topics don't have 110 years of intervals and haldane rate changes over time we will only look at intervals 1:30 to maintain a fair comparison
`````{r}
dat<-subset(dat, x<=30)
checkall<- ddply(dat, .(topic), summarise,
length.haldests =length(y),
length.intervals =length(unique(x)))
```````

The function below needs the same length for each variable. Since topic lengths differ, we pull out a random selection of the minimum, 255. This results in the loss of a lot of data for longer topics, but the haldane gams will still be representative.
````{r}
dat2<-ddply(dat,.(topic),function(x) x[sample(nrow(x),255),])
`````

Subset to one group
````{r}
dat.sub<-subset(dat2, topic=="topic100")
max(dat.sub$x)
plot(dat.sub$x, dat.sub$y)
``````
#### fitting a spline

since the x (interval) variable, and the spline points along it are the same in all cases, we only specify one set of them.  We fit spline points at 10 year intervals (we have previously fitted the maximum allowed by the data) to describe how abs.hald.rate varies as a function of interval. This function forces the spline through the origin (0,0)
```{r}
s<-seq(from=0, to=30, length.out=10) # sequence for knots
knots <-data.frame(x=s) # sets number of knots
sm <- smoothCon(s(x,k=10,bs="cr"),dat.sub, knots=knots)[[1]]
X <- sm$X[,-1] ## spline basis 
S <- sm$S[[1]][-1,-1] ## spline penalty 
off <- dat.sub$y*0 + 0 ## offset term to force curve through (0, 0)
````
iterate the model across all groups
```{r}
modelList2<-dlply(dat2, "topic", function(topic) gam(y ~ X - 1 + offset(off),paraPen=list(X=list(S)), se.fit=TRUE, data=topic))
````
make a new set of data to predict the splines.  We'll predict them at half-yearly intervals
```{r}
x<-seq(from=0, to=30, length=61) #make a new set of indepdendnt variables
newData <- data.frame(x) #put in a dataframe
names(newData) <- c("x")
````
predict the constrained gam
```{r}
Xp <- PredictMat(sm,data.frame(x=x))[,-1] #make a set of spline points from those new independent variables
pred<-ldply(modelList2, function(model) {
  data.frame(newData, predict(model,list(X=Xp,off=rep(0,61)), se.fit=TRUE)) #predict
})	
````
get haldane rates and plot
```{r}
pred$hald.rate<-pred$fit/pred$x
````

plot of abs.hald.rate gams
````{r}
ggplot(pred,aes(x=x,y=fit,group=as.factor(topic))) + geom_path(colour="red", alpha=0.5) + 
theme(legend.position="none") + xlab('interval') + ylab('fit')
````

plot of haldane rate gams
`````{r}
ggplot(pred,aes(x=x,y=hald.rate,group=as.factor(topic))) + geom_path(colour="red", alpha=1, size=0.1) + 
theme(legend.position="none") + xlab('interval') + ylab('haldane rate')
````

get topic labels and merge with pred. Remember: the topic labels of the gams are haldtopics
```{r}
d<-fread("topiclabels_4.csv", header=TRUE)
d<-as.data.frame(d)
d<-d[c("topic_number","use", "topic_discipline", "topic_majortaxon","topic_label")]
d$topic_number<-paste("topic", d$topic_number, sep="")
e<-fread("topicnames.csv", header=TRUE)
e<-as.data.frame(e)
e<-e[c("originaltopic","haldtopic", "order")]
e<-merge(e,d, by.x="originaltopic", by.y="topic_number")
pred<-merge(pred,e, by.x="topic", by.y="haldtopic")
`````
look at the 10 year rates
```{r}
int10<-subset(pred, x==10)
int10 <- int10[order(-int10$hald.rate),]
`````
look at the 25 year rates
```{r}
int25<-subset(pred, x==25)
int25 <- int25[order(-int25$hald.rate),]
hist(int25$hald.rate)
`````

```{r}
head(int10)
head(int25)
rates<-int10[c(1,2,3,5,6,11)]
ratesb<-int25[c(1,2,3,5,6,11)]
rates<-rbind(rates, ratesb)
rates$haldtopic<-rates$topic
z2<-z[c(2,3)]
rates<-merge(rates, z, by.x="haldtopic")
rates<-rates%>%
  filter(topic_label=="macroecology 1" | topic_label=="conservation" | topic_label=="conservation" | topic_label=="phylogenetics molecular" | topic_label=="climate change" | topic_label=="molecular population genetics" | topic_label=="optimal foraging" | topic_label=="phylogenetics" | topic_label == "natural selection & adaptation" | topic_label =="parental care" | topic_label =="social behaviour 1" | topic_label=="population genetics 1" | topic_label =="speciation" | topic_label=="community ecology" | topic_label=="macroecology 2")
rates<-arrange(rates, topic_label)
write.csv(rates, "Haldane rates & start years.csv")
```

```{r}
write.csv(pred, file="EEdata_05_haldrates.csv", row.names=FALSE)
```


