---
title: "rate calculator for EE data papers"
output: html_document
---
#### libraries
This is the haldane rate calculator for the EE data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#### libraries
````{r}
library(ggplot2)
library(scales)
library(plyr)
library(reshape2)
```````
#### data
`````{r}
taDF <- read.csv('EEpapers_topics_byyear_05.csv')
``````

`````{r}
taDF$year<-as.numeric(taDF$year)
``````
check years
````{r}
min(taDF$year)
max(taDF$year)
`````
Since the haldane calculator relies on SD and means, we should eliminate years with absurdly low sample sizes. We subset between 1900 and 2010.
```{r}
taDF<-subset(taDF, year>=1900 & year<=2010)
`````
Well, even so, we have on 113 papers in 1900

The calculator needs sequential topics. So we are going to use the haldtopics.  It's clutzy, but better than breaking the calculator function.
```{r}
taDF$topic<-taDF$haldtopic
````
This is a function that wraps the whole thing.
```{r}
fGetMidyearIntervalDiff <- function(taDF){
  aLen <- length(unique(taDF$topic))  # This gets the number of series
  bLen <- length(taDF$topic[taDF$topic=="topic1"]) # This gets the number of years within a series
  
  lResults <- data.frame() # This sets up a data frame
  
 # This  
  for (i in 1:aLen){ 
    print(i)
    aTopicName <- paste0("topic",i)
     if(i >= 99){
      aTopicName <- paste0("topic",i+1)
     }
    onetopic<-subset(taDF, topic==aTopicName)
    iter<-1:bLen   
    
    
    interval<-c() #an empty vector to put the results in
    
    for (j in iter){
      interval<-c(interval,(onetopic$year[j+iter]-onetopic$year[j]))
    }
    
    midyear<-c() #an empty vector to put the results in
    for (j in iter){
      midyear<-c(midyear,(onetopic$year[j]+((onetopic$year[j+iter]-onetopic$year[j])/2)))
    }
    
    diff<-c()
    for (j in iter){
      diff<-c(diff,(onetopic$proportion_present[j+iter]-onetopic$proportion_present[j]))
    }
    
    SDp<-c() #
	for (j in iter){
	SDp<-c(SDp,((onetopic$sd_proportion_present[j+iter]*(onetopic$N[j+iter]-1)+ onetopic$sd_proportion_present[j]*(onetopic$N[j]-1))/(onetopic$N[j+iter]+onetopic$N[j]-2)))
}

    
    # put these in a dataframe
    res<-as.data.frame(midyear)
    res$interval<-interval
    res$diff<-diff
    res$SDp<-SDp
    res$topic <- onetopic$topic
    lResults <- rbind.data.frame(lResults,res)
  }
  return(lResults)
}
````
run haldane calculator
````{r}
lResults <- fGetMidyearIntervalDiff(taDF)
````
additional manipulations. 
```{r}
lResults$abs.diff<-abs(lResults$diff) # get absolute differences
lResults$abs.hald.num<-lResults$abs.diff/lResults$SDp # get haldane numerator of abs.differences
lResults$hald.rate<-lResults$abs.hald.num/lResults$interval
```
sorts it nicely by run, interval and midyear
````{r}
lResults<- lResults[order(lResults$topic, lResults$interval, lResults$midyear),] 
````
get rid of NAs
````{r}
lResults<-lResults[complete.cases(lResults),]
````

check date ranges
````{r}
s <- ddply(lResults, .(topic), summarise,
			   min =min(midyear),
			   max = max(midyear),
			   minint = min(interval),
			   maxint = max(interval))
s <- s[order(s$min),] 
write_csv(s, "Topic_starts.csv")
````
So, some topics only begin at 1912 or 1931.


```{r}
write.csv(lResults, file="EEpapers_05_res.csv", row.names=FALSE)	
````
